%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%         PREAMBLE         %%%%%%%%%%%%

%% Default settings format a final copy (12pt font, single-sided,
%% double-spaced, normal margins, single-spaced notes).  For a rough
%% copy (10pt font, double-sided, double-spaced, normal margins, with
%% the word "DRAFT" printed at each corner of every page), use the
%% `draft' option.  The default line spacing can be changed with one
%% of the following options: `singlespaced', `oneandahalfspaced', or
%% `doublespaced'.  The notes are always single-spaced by default, but
%% can be made to have the same spacing as the rest of the document by
%% using the option `spacednotes'.  The size of the margins can be
%% changed with one of the following options: `narrowmargins' (1 1/4"
%% left, 3/4" others), `normalmargins' (1 1/4" left, 1" others),
%% `widemargins' (1 1/4" all), `extrawidemargins' (1 1/2" all).  Any
%% other standard option for the `report' document class can be used
%% to override the default or draft settings.

%% ***   Add any desired options.   ***
\documentclass[letterpaper]{ut-thesis} % For Single Sided
%\documentclass[letterpaper,twoside,10pt]{ut-thesis} % For Double Sided
%% ***   Add \usepackage declarations here.   ***
% Setting up graphics
\usepackage{ifpdf}

\ifpdf
\usepackage[pdftex]{graphicx}
\usepackage[protrusion=true,expansion=true]{microtype}
\else
\usepackage{graphicx}
\fi

\usepackage{longtable}
\usepackage{amsmath,amssymb,multirow,dcolumn,fancyhdr,charter,graphicx,color}
\usepackage{xspace}
\usepackage[hyphens]{url}

% Symbol Stuff
\usepackage{latexsym}%
\usepackage{amssymb}%

% Bibliography Stuff 
\usepackage{natbib}
\setcitestyle{aysep={},yysep={;}}

% AAS Tex Stuff
\usepackage{aastex_hack}
\usepackage{deluxetable}
\usepackage[labelsep=colon,small]{caption}

% Fonts
% Using Better Fonts
\usepackage[sc]{mathpazo}
\usepackage[T1]{fontenc}

% Paragraph Spacing Issue:
\setlength{\parskip}{0cm plus1mm minus1mm}

%commenting
\def\memohr#1{\color{blue}$HR[${\bf #1}$]$ \color{black}}
\def\memoas#1{\color{red}$AS[${\bf #1}$]$ \color{black}}

%% The line spacing of the document should be specified using one of
%% the document options given above, but if you need a line spacing
%% that is not provided by the options, you can override the default
%% line spacing for the entire document with the command
%%   `\linespacing{...}'.
%% Note that in order to get the correct appearance, the argument to
%% `\linespacing' must be equal to 1/3 + 2/3 times the desired line
%% spacing (for example, single-spaced = \linespacing{1},
%%                        1 1/2-spaced = \linespacing{1.33}, and
%%                       double-spaced = \linespacing{1.66}).

%% ***   Uncomment and fill in a value, if needed.    ***
%% ***   REMEMBER: You should NOT need to use this.  Use one of   ***
%% ***   the document class options mentionned above instead.     ***
%\linespacing{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                  %%
%%                  ***   I M P O R T A N T   ***                   %%
%%                                                                  %%
%%  Fill in the following fields with the required information:     %%
%%   - \degree{...}       name of the degree obtained               %%
%%   - \department{...}   name of the graduate department           %%
%%   - \gradyear{...}     year of graduation                        %%
%%   - \author{...}       name of the author                        %%
%%   - \title{...}        title of the thesis                       %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% ***   Change this example to appropriate values.   ***
\degree{Doctor of Philosophy} \department{Astronomy \& Astrophysics}
\gradyear{2017} \author{Ari Silburt} \title{Statistics, Formation and Stability of Exoplanetary Systems}

%% ***   NOTE   ***
%% Put here all other formatting commands that belong in the preamble.


%% For example, to list only down to subsections in table of contents
%% (-1=part, 0=chapter, 1=section, 2=subsection, 3=subsubsection,
%%  4=paragraph, 5=subparagraph, 6=subsubparagraph).
%
\setcounter{tocdepth}{2}


%%%%%%%%%%%%      MAIN  DOCUMENT      %%%%%%%%%%%%

\begin{document}


%% ***   NOTE   ***
%% You should put all of your `\newcommand', `\newenvironment', and
%% `\newtheorem's (in other words, all the global definitions that
%% you will need throughout your thesis) in a separate file and use
%% "\input{filename}" to input it here.

\input{thesiscommands}

%% This sets the page style and numbering for preliminary sections.
\begin{preliminary}

%% This generates the title page from the information given above.
\maketitle

%% There should be NOTHING between the title page and abstract.

%% This generates the abstract page, with the line spacing adjusted
%% according to SGS guidelines.
\begin{abstract}
  
Over the past two decades scientists have detected thousands of exoplanet detections through a variety of methods, and collective properties of exoplanetary systems are now emerging. 
This thesis contributes to the exoplanet field by analyzing the statistics, formation and stability of exoplanetary systems.

The first part of this thesis conducts a statistical reconstruction of the radius and period distributions of \kep planets. 
Accounting for observation and detection biases, as well as measurement errors, we calculate the occurrence of planetary systems, including the prevalence of Earth-like planets.
This calculation is compared to related works, finding both similarities and differences. 
 
Second, the formation of \kep planets near mean motion resonance (MMR) is investigated.
In particular, 27 \kep systems near 2:1 MMR are analyzed to determine whether tides are a viable mechanism for transporting \kep planets from MMR.
We find that tides alone cannot transport near-resonant planets from exact 2:1 MMR to their observed locations, and other mechanisms must be invoked to explain their formation.

Third, a new hybrid integrator \hermes is presented, which is capable of simulating N-bodies undergoing close encounters.
\hermes is specifically designed for planets embedded in planetesimal disks, and includes an adaptive routine for optimizing the close encounter boundary to help maintain accuracy.
We find the performance of \hermes comparable to other popular hybrid integrators.
 
Fourth, the longterm stability of planetary systems is investigated using machine learning techniques.
Typical studies of longterm stability require thousands of realizations to acquire statistically rigorous results, which can take weeks or months to perform. 
Here we find that a trained machine is capable of quickly and accurately classifying longterm planet stability. 
 
Finally, the planetary system HD155358, consisting of two Jovian-sized planets near 2:1 MMR, is investigated using previously collected radial velocity data. 
New orbital parameters are derived using a Bayesian framework, and from this we find a high likelihood that the planets are in MMR.
In addition, constraints are placed on the formation history of HD155358, and our best-fit parameters land firmly in dynamically stable regions of parameter space. 
 
\end{abstract}
\cleardoublepage
%% Anything placed between the abstract and table of contents will
%% appear on a separate page since the abstract ends with \newpage
%% and the table of contents starts with \clearpage.

%% This generates a "dedication" section, if needed.
%% (uncomment to have it appear in the document)
%%\begin{dedication}
%%My Dedication
%%\end{dedication}

% A better opening Quote Environment

\vspace*{\fill}
\begin{center}
\begin{minipage}[c]{4.75in}
 "At the end of the day people won't remember what you said or did, they will remember how you made them feel." \vspace{1em}
 
\hfill \emph{-Maya Angelou}


\end{minipage}
\end{center}
\vspace*{\fill}

\cleardoublepage
%\newpage
%% The `dedication' and `acknowledgements' sections do not create new
%% pages so if you want the two sections to appear on separate pages,
%% you should put an explicit \newpage between them.

%% This generates an "acknowledgements" section, if needed.
%% (uncomment to have it appear in the document)
\begin{acknowledgements}
%Hanno + Hanno's razor
First and foremost I would like to thank my supervisor, Hanno Rein. 
Looking back to the beginning of my doctorate program, I realize now that, at the time I did not really know how to conduct quality research or even how to ask the right questions.
Through your hard work in answering all my endless questions, guiding me in fruitful research directions, and teaching me the skills required to succeed, I now feel more confident in my research abilities than I ever thought possible. 
You are the greatest academic mentor that I've had in my life.

%Dan
Next, I would like to thank Dan Tamayo, who is one of the greatest humans I know. 
Compassionate, intelligent, patient and humble, you have done so much for me as well as the exoplanet community. 
From hosting machine learning workshops and dynamics discussion groups, to late-night collaborations for impending postdoc deadlines, to answering my endless stream of incompetent questions, you are a role model for the kind of person I would like to be. 

%Eric, Yanqin
Eric Gaidos and Yanqin Wu, you have been invaluable throughout my doctorate, and I owe you a world of thanks. 
Our publication was my first real introduction to research, and it is only through your patience, intelligence and kindness that I was able to see the job through to the end. 
I'm grateful that you trusted in my abilities, wrote endless reference letters for me, and continually supported me throughout my doctorate.

%Joey, Bruce, Barbara.
My family has always been endlessly supportive of who I am, and especially so during my doctorate.
To my parents, Bruce and Barbara, I remember as a young child looking up at the stars with you in our backyard in Guelph, contemplating the biggest questions about the universe and igniting my passion for science.
This doctorate thesis is a realization of that dream; a flame that never burned out through all these years. 
To my brother Joey, thank you for the countless discussions about science, machine learning, life, etc., grounding me during stressful times, and the fun activities we've done outside the academic realm (Oma's, Age of Empires, etc.). 
I've cherished the time we've lived together in Toronto, and will miss you when I leave. 

%Max, Charles & Yev, Lisa, 
Many colleagues at the department have made my time in Toronto very special. 
In particular, Max Millar-Blanchaer who has been an incredible friend, role model, and advisor, Yevgeni Kissin and Chenchong Zhu who have been my machine learning buddies over the last few years, and Lisa Esteves who I've had awesome chills, exoplanet discussions and philosophical conversations with during my time at U of T.
I'd also like to thank my graduation year -- APLEJIBS -- who I've spent five years with in the trenches, and I'm really glad that we made the extra effort to hang out in our final year together. 

%CPS
I would like to thank the Centre for Planetary Science (CPS) in Scarborough, Canada, as well as all its members, who have been an integral part of my development as a researcher. 
Through machine learning workshops, CPS planet days, lunchtime discussions and more, I have become a much more competent researcher than I otherwise would have ever been had CPS not been in my life. 

%Money sources - U of T, NSERC CGS M, PGS D.
Lastly, I would like to thank the institutions and people that have funded me throughout this doctoral degree. 
In particular, the Natural Sciences and Engineering Research Council of Canada (NSERC) for CGS M and PGS D grants, the University of Toronto Reinhardt fund for travel reimbursement, the Walter C Sumner Memorial Fellowship Foundation, the University of Toronto Department of Astronomy \& Astrophysics, and my supervisor Hanno Rein for providing me the necessary funds and resources to complete my research. 

\end{acknowledgements}

%% This generates the Table of Contents (on a separate page).
\tableofcontents

%% This generates the List of Tables (on a separate page), if needed.
%% (uncomment to have it appear in the document)
\listoftables

%% This generates the List of Figures (on a separate page), if needed.
%% (uncomment to have it appear in the document)
\listoffigures

%% End of the preliminary sections: reset page style and numbering.
\end{preliminary}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  Put your Chapters here; the easiest way to do this is to keep   %%
%%  each chapter in a separate file and `\include' all the files    %%
%%  right here.  Note that each chapter file should start with the  %%
%%  line "\chapter{ChapterName}".  Note that using `\include'       %%
%%  instead of `\input' makes each chapter start on a new page.     %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% ***   Include chapter files here.   ***

\include{intro/intro}
\include{chap2/chap2}
\include{chap3/chap3}
\include{chap4/chap4}
\include{chap5/chap5}
\include{chap6/chap6}
\include{conc/conc}

%% This adds a line for the Bibliography in the Table of Contents.
\addcontentsline{toc}{chapter}{Bibliography}

%% ***   Set the bibliography style.   ***
%% (change according to your preference)
\bibliographystyle{apj}
%% ***   Set the bibliography file.   ***
%% ("thesis.bib" by default; change if needed)
\bibliography{thesis}

%% ***   NOTE   ***
%% If you don't use bibliography files, comment out the previous line
%% and use \begin{thebibliography}...\end{thebibliography}.  (In that
%% case, you should probably put the bibliography in a separate file
%% and `\include' or `\input' it here).

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  End of UT-THESIS.TEX
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
